Thermal performance of fish is explained by an interplay between physiology, behaviour, and ecology

### 2019-01-16 MS_AMNat.Rnw
revisions

The oxygen budget $P_{0_2}(w,T)$ (or aerobic scope) follows a similar form to the mass budget:
\begin{align}
  P_{0_2}(w,T) &= S_{O_2}(w,T) - D_{O_2}(w,T) \\
        &= S_{O_2}c_{O_2}(T) w^n - \omega c(T) \left( \beta f(w,T) h w^q + k w^n + k_a w \right).
\end{align}
Demand ($D_{O_2}(w,T)$) is the sum of oxygen used for all metabolic processes in Eq.~\ref{eq:P} (except assimilation losses), with $\omega$ being amount of oxygen required per mass.  The oxygen supply ($S_{O_2}c_{O_2}(T)$) scales with body weight as $w^n$ multiplied by a flexible dome-shaped function that can emulate both a dome-shaped maximum oxygen supply (MOS) as well as a MOS that increase continuously up to a lethal temperature (fig./~\ref{fig:O2_fig_used}). The maximum oxygen consumption is the oxygen consumption during maximal activity level that can be sustained over some time, and is usually termed the maximum metabolic rate (MMR). Although the MMR is often used synonymously with both maximum oxygen supply and demand, in some species the maximal oxygen consumption ($D^{max}_{O_2}(T)$) is not reached at maximum activity levels, but rather during digestion (\citealt{priede_metabolic_1985}). We use MMR and oxygen supply interchangeably, although maximum oxygen demand (i.e., at full activity) may be lower than the MMR. We assume that oxygen supply is temperature dependent and follows a flexible dome-shaped function (\citealt{gnauck2013freshwater, lefrancois_influence_2003}):

@@

Here $\lambda(T)$ specifies the temperature dependency of $O_2$ supply, whereas the $s_{O_2}(T)$ term describes the saturation dependence on ambient $O_2$ at temperature $T$ ($O_2(T)$). At constant temperature $T$, oxygen supply is a function of ambient oxygen and is assumed to follow a saturating function (\cite[e.g.,][]{lefrancois_influence_2003}). We specify $P_{50}$ as the point where oxygen supply has dropped by 50\% relative to the saturation level $\lambda(T)$, and $P_{\text{crit}}$ is the ambient concentration at which oxygen supply ceases. Ambient oxygen levels are assumed to decline with temperature as $l*\exp(-0.01851*(T-5)$, with $l$ the ambient oxygen concentration at 5\degree C. To specify $\lambda(T)$, we use $T_{\text{max}}$ the lethal temperature for the species, $T_{\text{opt}}$ the temperature at which oxygen supply is maximised; $\eta$ determines the width of the dome-shape, and $\zeta$ its height. Note that we can emulate an oxygen supply (and hence maximum metabolic rate; MMR) that increases up to the lethal temperature by setting the temperature for maximum oxygen delivery close to the lethal temperature (fig./~\ref{fig:O2_fig}).

\input{O2_fig}

@@

The slope of maturation reaction norms is evolutionarily determined by the strength of the co-variation between growth and mortality for a given population in a given environment. Strongly positive co-variation leads to the relatively flat reaction norms observed for most fish populations (\citealt{marty_impact_2011}). This covariation is probably the consequence of good growth conditions (e.g., from increased prey density) altering the baseline mortality $\rho$ and the risk of foraging $\eta$ (e.g., by attracting predators). We did not explicitly model these interactions here (aside from the dependence of mortality on $\tau$), but rather assumed that reaction norms evolved over regimes of relatively stable temperature and growth variation in the past. Consequently, we assumed that the evolved slope of the reaction norm is a fixed trait over the time-scales considered here for a particular species or population, and for simplicity and generality we assume a flat reaction norm (i.e., allocation to reproduction is a function of size only; fig./~\ref{fig:MRN}), although sloping reaction norms can be formulated and used in our framework (see appendix \ref{sec:MRN}). The intercept of this reaction norm was found numerically by maximising fitness ($R_0$, see below) at the reference temperature for our simulations (15\degree C), and this intercept was assumed fixed as temperatures change.

@@

<< setup,echo=FALSE,include=F>>=
knitr::opts_chunk$set(echo = F, message = F,cache=T,autodep = T)

@@

lab1 <- as_labeller(c(`M strategy Domed`  ='a',
                   `P strategy Domed` = 'c'))

lab2 <- as_labeller(c(`M strategy Domed`  ='b',
                   `P strategy Domed` = 'd'))

>>>>

The oxygen budget $P_{0_2}(w,T)$ (or aerobic scope) follows a similar form to the mass budget:
\begin{align}
  P_{0_2}(w,T) &= S_{O_2}(w,T) - D_{O_2}(w,T) \\
        &= S_{O_2}c_{O_2}(T) w^n - \omega c(T) \left( \beta f(w,T) h w^q + k w^n + k_a w \right).
\end{align}
Demand ($D_{O_2}(w,T)$) is the sum of oxygen used for all metabolic processes in Eq.~\ref{eq:P} (except assimilation losses), with $\omega$ being amount of oxygen required per mass.  The oxygen supply ($S_{O_2}c_{O_2}(T)$) scales with body weight as $w^n$ multiplied by a flexible dome-shaped function that can emulate both a dome-shaped maximum oxygen supply (MOS) as well as a MOS that increase continuously up to a lethal temperature (fig.~\ref{fig:O2_fig_used}). The maximum oxygen consumption is the oxygen consumption during maximal activity level that can be sustained over some time, and is usually termed the maximum metabolic rate (MMR). Although the MMR is often used synonymously with both maximum oxygen supply and demand, in some species the maximal oxygen consumption ($D^{max}_{O_2}(T)$) is not reached at maximum activity levels, but rather during digestion (\citealt{priede_metabolic_1985}). We use MMR and oxygen supply interchangeably, although maximum oxygen demand (i.e., at full activity) may be lower than the MMR. We assume that oxygen supply is temperature dependent and follows a flexible dome-shaped function (\citealt{gnauck2013freshwater, lefrancois_influence_2003}):

@@

Here $\lambda(T)$ specifies the temperature dependency of $O_2$ supply, whereas the $s_{O_2}(T)$ term describes the saturation dependence on ambient $O_2$ at temperature $T$ ($O_2(T)$). At constant temperature $T$, oxygen supply is a function of ambient oxygen and is assumed to follow a saturating function (\cite[e.g.,][]{lefrancois_influence_2003}). We specify $P_{50}$ as the point where oxygen supply has dropped by 50\% relative to the saturation level $\lambda(T)$, and $P_{\text{crit}}$ is the ambient concentration at which oxygen supply ceases. Ambient oxygen levels are assumed to decline with temperature as $l*\exp(-0.01851*(T-5)$, with $l$ the ambient oxygen concentration at 5\degree C. To specify $\lambda(T)$, we use $T_{\text{max}}$ the lethal temperature for the species, $T_{\text{opt}}$ the temperature at which oxygen supply is maximised; $\eta$ determines the width of the dome-shape, and $\zeta$ its height. Note that we can emulate an oxygen supply (and hence maximum metabolic rate; MMR) that increases up to the lethal temperature by setting the temperature for maximum oxygen delivery close to the lethal temperature (fig.~\ref{fig:O2_fig_used}).

@@

The slope of maturation reaction norms is evolutionarily determined by the strength of the co-variation between growth and mortality for a given population in a given environment. Strongly positive co-variation leads to the relatively flat reaction norms observed for most fish populations (\citealt{marty_impact_2011}). This covariation is probably the consequence of good growth conditions (e.g., from increased prey density) altering the baseline mortality $\rho$ and the risk of foraging $\eta$ (e.g., by attracting predators). We did not explicitly model these interactions here (aside from the dependence of mortality on $\tau$), but rather assumed that reaction norms evolved over regimes of relatively stable temperature and growth variation in the past. Consequently, we assumed that the evolved slope of the reaction norm is a fixed trait over the time-scales considered here for a particular species or population, and for simplicity and generality we assume a flat reaction norm (i.e., allocation to reproduction is a function of size only; fig.~\ref{fig:MRN}), although sloping reaction norms can be formulated and used in our framework (see appendix \ref{sec:MRN}). The intercept of this reaction norm was found numerically by maximising fitness ($R_0$, see below) at the reference temperature for our simulations (15\degree C), and this intercept was assumed fixed as temperatures change.

@@

To implement this axis, we used the result that species with a more
active, production oriented life-history (e.g., predatory pelagic
fish) have a higher standard metabolism and lower weight scaling of
metabolic costs
(\citealt{priede_metabolic_1985,killen_intraspecific_2010}). We assumed
that higher standard metabolism is due to increased digestive capacity
(i.e., is used for gut maintenance), though high muscle mass and a
larger heart will also contribute to higher standard metabolism in
active species (\citealt{priede_metabolic_1985}). In practice, we assumed
that approximately 50\% of the standard metabolic cost is due to gut
maintenance, such that a doubling of the maximum ingestion leads to a
50\% increase in standard metabolic cost. We further assumed that such active species have a less effective refuge from predators and therefore have a higher constant mortality, but lower mortality related to activity (i.e., $M_{s}(w) = (0.1+6\tau) w^{q-1}$ and $M_{f}(w) = (1+\tau) w^{q-1}$). Exact parameter values for these trait scenarios are given in Table \ref{tab:parameters}. Together, these assumed trait differences lead to very different ecological and bio-energetic responses of slow- and fast strategists (fig.~\ref{fig:scenarios}), with $\tau_{\text{opt}}$ found at lower activity levels for slow-strategists, as high activity induces exceedingly high mortality and decreasing energy efficiency (i.e., available energy relative to food intake) at high activity levels. A slower increase in available energy and M with $\tau$ for fast strategists leads to a higher $\tau_{\text{opt}}$.

We further contrasted species with oxygen limitation at high temperatures (i.e., species with a uni-modal metabolic scope) with species that do not experience oxygen limitation at high temperatures (at least not up to a lethal temperature, where death may be induced by sudden failure to deliver oxygen to vital organs, or failure of bio-chemical pathways at high temperature \citealt{salin_inadequate_2016,iftikar_mitochondria_2013}). In practice, this was achieved as described above by setting the maximum oxygen delivery close to the lethal temperature (fig.~\ref{fig:O2_fig_used}). Note that, although we assume here that limitations over the temperature range are due to oxygen availibility, other limiting mechanisms, such as the respiratory control ratio (\citealt{salin_inadequate_2016,iftikar_mitochondria_2013}) may determine upper limits to activity over some or all of a species temperature range. However the over-all mechanism would be the same to the one assumed here, with different untis (e.g., ATP instead of $O_2$).

Increasing metabolic demands with temperature lead species to adjust activity levels in order to optimise energy gains relative to mortality risk (fig.~\ref{fig:activity}). This adjustment is especially pronounced in slow strategy species, for which the over-all activity level is markedly lower and which show a consistent increase in activity over-all sizes for the simulated life-history (fig.~\ref{fig:size-act}). A similar increase in activity is observed for small fast strategy individuals (e.g., post-larval) for which even initial activity levels are very high (fig.~\ref{fig:size-act}). For these individuals, the increasing activity and metabolic demands lead to an active metabolic rate that is close to their MMR. For all other sizes across the two trait scenarios, oxygen is only limiting to activity at the extremes of the simulated temperature range (fig.~\ref{fig:activity}), and only for species with a dome-shaped MOS with respect to temperature. For species with a rising MOS with temperature, oxygen is not limiting (fig.~\ref{fig:activity-rising}). However, larger fast strategy individuals are predicted to show a slightly dome-shaped adjustment in activity levels at intermediate sizes, and a slightly decreasing activity level in response to temperature at large sizes, despite available aerobic scope for activity. This adjsutment is a function of metabolic activity costs assumed here - at smaller metabolic costs per activity, activity adjustment are strictly increasing (fig.~\ref{fig:size-act_low}).

Temperature and metabolic demand-driven adjustments to the activity level lead to substantial changes in performance related metrics in both trait scenarios (fig.~\ref{fig:feeding}). For slow strategy species, the increases in activity levels lead to relatively stable feeding levels, but a substantial increase in mortality coupled with a slow increase in available energy leads to an over-all decline in the ratio of $P$ to $M$. Available energy shows a dome-shaped response to temperature in slow strategists, and is maximised at relatively high temperatures. However, it is limited by oxygen availability only at high temperatures in species with a dome-shaped MOS. Production efficiency follows a near opposite trend due to the relatively flat response in the feeding level $f$, but temperature driven increases in maximum consumption.

For fast strategists, the relatively modest response in activity levels at all but the smallest sizes leads to a decline in feeding levels, which causes a largely dome-shaped response of available energy and growth efficiency to warmer temperatures (fig.~\ref{fig:feeding}). Again, production efficiency peaks at relatively low temperatures, but in this scenario available energy $P$ peaks at much lower temperatures. Given the relatively flat mortality levels, the ratio of $P/M$ largely follows the trend in $P$.

Simulated growth curves illustrate the ontogenetic consequences of increasing temperatures (fig.~\ref{fig:norms}). For both trait scenarios, fastest growth occurred at relatively high temperatures, with declining growth for oxygen limited species at the highest temperatures (fig.~\ref{fig:norms_O2}). This can be explained by ontogenetic shifts in temperature optima for growth (fig.~\ref{fig:size-rates}): for small individuals, available energy and growth consistently peak at high temperatures, but this peak rapidly moves to lower temperatures as individuals in either trait-scenario grow. For large individuals, growth is optimised at relatively low temperatures, leading to larger asymptotic size at lower temperatures.

Resource availability strongly modulates this growth response to temperature: low resource availability leads to strong differences in asymptotic size, whereas high food availability leads to fast growth and larger asymptotic length at high temperatures (figs.~\ref{fig:norms},\ref{fig:infs}). In addition, in very resource poor conditions, individuals may not grow to reproductive size in our scenario of a flat MRN. Over-all growth responses to temperature are not strongly affected by the assumed slope of the MRN (figs.~\ref{fig:norms_neg},\ref{fig:norms_pos}), although a negatively sloped MRN does ensure maturation in low resource environments.

Over-all fitness consequences mirror trends in the ratio of $P/M$ (fig.~\ref{fig:fit}a), which can be seen as a short sighted approximation to over-all fitness optimisation (\citealt{sainmont_effective_2015}). With increasing temperatures, fitness declines at our basic parameter settings, in opposition to growth and aerobic scope. At low temperatures, fitness is limited by aerobic scope, with the magnitude determined by the extend of doming in aerobic scope. Note that this limitation through the aerobic scope appears at higher temperatures than apparent from fig.~\ref{fig:activity}, reflecting stronger limitation of oxygen on growth during early life (fig.~\ref{fig:size-rates}). Fitness trends with temperature are strongly dependent on metabolic costs of activity (fig.~\ref{fig:fit}b), and changing the activity cost to lower values attenuates the decline in fitness with temperature for slow strategy species, and moves the fitness optimum to higher temperatures for fast strategy species. Similarly, increased food availibility leads to changes in fitness optima, with higher optimum temperatures, especially for fast strategists, and a slower decline of fitness with temperature.

Our quantitative thermal impact model generalises existing eco-physiological models for particular species and stocks (\citealt{hufnagl_physiological_2011, holt_climate_2014,holt_climate_2015}), and allows to develop a more nuanced understanding of interactions between temperature, oxygen limitation and ecology for species with varying traits. In line with the conceptual framework of Fry's aerobic scope and the OCLTT, our model suggests that oxygen limitation can be a potentially important ecological driver, especially at extreme temperatures for species with declining MOSs in such temperature regimes. At the onset of this limitation, ecological parameters change drastically, and growth, mortality are strongly impacted. This limitation closely mimics limitations seen in wild fish (\citealt{myrick_temperature_2000}), and is in line with observations that fish often seek specific water temperatures to optimise metabolic function (\cite[e.g., ][]{armstrong_diel_2013,claireaux_physiology_1995}). Fitness, however, appears to be limited through the metabolic scope primarily via limitations at temperature extremes and impact on particular life-history stages. For instance, oxygen limitation is a more severe constraint for small individuals (fig.~\ref{fig:size-rates}), and thereby can limit growth performance early in life, impacting over-all fitness. Furthermore, changes in environmental oxygen supply, if beyond an organisms ability to compensate via passive or active compensation mechanisms, will induce an over-all lower aerobic scope and lead to an earlier onset of oxygen limitation, but such scenarios do not change the qualitative predictions from our model.

\titleformat{\section}{\normalfont\bfseries\sffamily}{APPENDIX \thesection:}{1em}{}

A flat maturation reaction norm (MRN) leads to a logistic allocation to reproduction that is only dependent on size and independent of age (fig.~\ref{fig:MRN}).

\bibitem[{Andersen et~al.(2009)Andersen, Farnsworth, Pedersen, Gislason,
  and Beyer}]{andersen_how_2009}
Andersen, K.~H., K.~D. Farnsworth, M.~Pedersen, H.~Gislason, and J.~E. Beyer.
  2009.
\newblock How community ecology links natural mortality, growth, and production
  of fish populations.
\newblock {ICES} Journal of Marine Science: Journal du Conseil 66:1978--1984.

\bibitem[{Angilletta et~al.(2004)Angilletta, Steury, and
  Sears}]{angilletta_temperature_2004}
Angilletta, M.~J., T.~D. Steury, and M.~W. Sears. 2004.
\newblock Temperature, growth rate, and body size in ectotherms: Fitting pieces
  of a life-history puzzle.
\newblock Integrative and Comparative Biology 44:498--509.

\bibitem[{Armstrong et~al.(2013)Armstrong, Schindler, Ruff, Brooks, Bentley,
  and Torgersen}]{armstrong_diel_2013}
Armstrong, J.~B., D.~E. Schindler, C.~P. Ruff, G.~T. Brooks, K.~E. Bentley, and
  C.~E. Torgersen. 2013.
\newblock Diel horizontal migration in streams: juvenile fish exploit spatial
  heterogeneity in thermal and trophic resources.
\newblock Ecology 94:2066--2075.

\bibitem[{Atkinson(1994)}]{atkinson_temperature_1994}
Atkinson, D. 1994.
\newblock Temperature and organism size: a biological law for ectotherms?
\newblock Advances in ecological research 25:1--1.

\bibitem[{Brander et~al.(2013)Brander, Neuheimer, Andersen, and
  Hartvig}]{brander_overconfidence_2013}
Brander, K., A.~Neuheimer, K.~H. Andersen, and M.~Hartvig. 2013.
\newblock Overconfidence in model projections.
\newblock {ICES} Journal of Marine Science 70:1065--1068.

\bibitem[{Brett(1965)}]{brett_1965_relation}
Brett, J. 1965.
\newblock The relation of size to rate of oxygen consumption and sustained
  swimming speed of sockeye salmon (oncorhynchus nerka).
\newblock Journal of the Fisheries Board of Canada 22:1491--1501.

\bibitem[{Brown et~al.(1989)Brown, Pepin, Methven, and
  Somerton}]{brown_feeding_1989}
Brown, J.~A., P.~Pepin, D.~A. Methven, and D.~C. Somerton. 1989.
\newblock The feeding, growth and behaviour of juvenile cod, gadus morhua l.,
  in cold environments.
\newblock Journal of Fish Biology 35:373--380.

\bibitem[{Brown et~al.(2004)Brown, Gillooly, Allen, Savage, and
  West}]{brown_toward_2004}
Brown, J.~H., J.~F. Gillooly, A.~P. Allen, V.~M. Savage, and G.~B. West. 2004.
\newblock Toward a metabolic theory of ecology.
\newblock Ecology 85:1771--1789.

\bibitem[{Charnov et~al.(2013)Charnov, Gislason, and
  Pope}]{charnov_evolutionary_2013}
Charnov, E.~L., H.~Gislason, and J.~G. Pope. 2013.
\newblock Evolutionary assembly rules for fish life histories.
\newblock Fish and Fisheries 14:213--224.

\bibitem[{Cheung et~al.(2013)Cheung, Sarmiento, Dunne, Frölicher, Lam,
  Palomares, Watson, and Pauly}]{cheung_shrinking_2013}
Cheung, W.~W., J.~L. Sarmiento, J.~Dunne, T.~L. Frölicher, V.~W. Lam, M.~D.
  Palomares, R.~Watson, and D.~Pauly. 2013.
\newblock Shrinking of fishes exacerbates impacts of global ocean changes on
  marine ecosystems.
\newblock Nature Climate Change 3:254--258.

\bibitem[{Claireaux and Lefrançois(2007)}]{claireaux_linking_2007}
Claireaux, G., and C.~Lefrançois. 2007.
\newblock Linking environmental variability and fish performance: integration
  through the concept of scope for activity.
\newblock Philosophical Transactions of the Royal Society B: Biological
  Sciences 362:2031--2041.

\bibitem[{Claireaux et~al.(1995)Claireaux, Webber, Kerr, and
  Boutilier}]{claireaux_physiology_1995}
Claireaux, G., D.~Webber, S.~Kerr, and R.~Boutilier. 1995.
\newblock Physiology and behaviour of free-swimming atlantic cod (gadus morhua)
  facing fluctuating temperature conditions.
\newblock Journal of Experimental Biology 198:49--60.

\bibitem[{Claireaux et~al.(2016-08-08)Claireaux, Webber, Lagardère, and
  Kerr}]{claireaux_influence_2000}
Claireaux, G., D.~M. Webber, J.-P. Lagardère, and S.~R. Kerr. 2016-08-08.
\newblock Influence of water temperature and oxygenation on the aerobic
  metabolic scope of atlantic cod (gadus morhua).
\newblock Journal of Sea Research 44:257--265.

\bibitem[{Deutsch et~al.(2015)Deutsch, Ferrel, Seibel, Pörtner, and
  Huey}]{deutsch_climate_2015}
Deutsch, C., A.~Ferrel, B.~Seibel, H.-O. Pörtner, and R.~B. Huey. 2015.
\newblock Climate change tightens a metabolic constraint on marine habitats.
\newblock Science 348:1132--1135.

\bibitem[{Dieckmann and Heino(2007)}]{dieckmann_probabilistic_2007}
Dieckmann, U., and M.~Heino. 2007.
\newblock Probabilistic maturation reaction norms: their history, strengths,
  and limitations.
\newblock Marine Ecology Progress Series 335:253--269.

\bibitem[{Englund et~al.(2011)Englund, Öhlund, Hein, and
  Diehl}]{englund_temperature_2011}
Englund, G., G.~Öhlund, C.~L. Hein, and S.~Diehl. 2011.
\newblock Temperature dependence of the functional response.
\newblock Ecology Letters 14:914--921.

\bibitem[{Fry(1947)}]{fry_effects_1947}
Fry. 1947.
\newblock Effects of the environment on animal activity .

\bibitem[{Glazier(2009)}]{glazier_activity_2009}
Glazier, D.~S. 2009.
\newblock Activity affects intraspecific body-size scaling of metabolic rate in
  ectothermic animals.
\newblock Journal of Comparative Physiology B 179:821--828.

\bibitem[{Gnauck and Stra{\v{s}}kraba(2013)}]{gnauck_2013_freshwater}
Gnauck, A., and M.~Stra{\v{s}}kraba. 2013.
\newblock Freshwater ecosystems: modelling and simulation, vol.~8.
\newblock Elsevier.

\bibitem[{Guiet et~al.(2016)Guiet, Aumont, Poggiale, and
  Maury}]{guiet_effects_2016}
Guiet, J., O.~Aumont, J.-C. Poggiale, and O.~Maury. 2016.
\newblock Effects of lower trophic level biomass and water temperature on fish
  communities: A modelling study.
\newblock Progress in Oceanography 146:22--37.

\bibitem[{Hartvig et~al.(2011)Hartvig, Andersen, and Beyer}]{hartvig_food_2011}
Hartvig, M., K.~H. Andersen, and J.~E. Beyer. 2011.
\newblock Food web framework for size-structured populations.
\newblock Journal of Theoretical Biology 272:113--122.

\bibitem[{Holt and Jorgensen(2014)}]{holt_climate_2014}
Holt, R.~E., and C.~Jorgensen. 2014.
\newblock Climate warming causes life-history evolution in a model for atlantic
  cod (gadus morhua).
\newblock Conservation Physiology 2.

\bibitem[{Holt and Jorgensen(2015)}]{holt_climate_2015}
Holt, R.~E., and C.~Jorgensen. 2015.
\newblock Climate change in fish: effects of respiratory constraints on optimal
  life history and behaviour.
\newblock Biology Letters 11:20141032--20141032.

\bibitem[{Hufnagl and Peck(2011)}]{hufnagl_physiological_2011}
Hufnagl, M., and M.~A. Peck. 2011.
\newblock Physiological individual-based modelling of larval atlantic herring
  (clupea harengus) foraging and growth: insights on climate-driven
  life-history scheduling.
\newblock {ICES} Journal of Marine Science 68:1170--1188.

\bibitem[{Iftikar and Hickey(2013)}]{iftikar_mitochondria_2013}
Iftikar, F.~I., and A.~J.~R. Hickey. 2013.
\newblock Do mitochondria limit hot fish hearts? understanding the role of
  mitochondrial function with heat stress in notolabrus celidotus.
\newblock {PLOS} {ONE} 8:e64120.

\bibitem[{Jeschke et~al.(2002)Jeschke, Kopp, and
  Tollrian}]{jeschke_predator_2002}
Jeschke, J.~M., M.~Kopp, and R.~Tollrian. 2002.
\newblock Predator functional responses: discriminating between handling and
  digesting prey.
\newblock Ecological Monographs 72:95--112.

\bibitem[{Jutfelt et~al.(2018)Jutfelt, Norin, Ern, Overgaard, Wang, {McKenzie},
  Lefevre, Nilsson, Metcalfe, and Hickey}]{jutfelt_oxygen-and_2018}
Jutfelt, F., T.~Norin, R.~Ern, J.~Overgaard, T.~Wang, D.~J. {McKenzie},
  S.~Lefevre, G.~E. Nilsson, N.~B. Metcalfe, and A.~J. Hickey. 2018.
\newblock Oxygen-and capacity-limited thermal tolerance: blurring ecology and
  physiology.
\newblock Journal of Experimental Biology 221:jeb169615.

\bibitem[{Killen et~al.(2010)Killen, Atkinson, and
  Glazier}]{killen_intraspecific_2010}
Killen, S.~S., D.~Atkinson, and D.~S. Glazier. 2010.
\newblock The intraspecific scaling of metabolic rate with body mass in fishes
  depends on lifestyle and temperature.
\newblock Ecology Letters 13:184--193.

\bibitem[{Lefevre(2016)}]{lefevre_are_2016}
Lefevre, S. 2016.
\newblock Are global warming and ocean acidification conspiring against marine
  ectotherms? a meta-analysis of the respiratory effects of elevated
  temperature, high {CO}2 and their interaction.
\newblock Conservation Physiology 4.

\bibitem[{Lefevre et~al.(2017)Lefevre, {McKenzie}, and
  Nilsson}]{lefevre_models_2017}
Lefevre, S., D.~J. {McKenzie}, and G.~E. Nilsson. 2017.
\newblock Models projecting the fate of fish populations under climate change
  need to be based on valid physiological mechanisms.
\newblock Global Change Biology .

\bibitem[{Lefevre et~al.(2018)Lefevre, McKenzie, and
  Nilsson}]{lefevre_modelling_2017}
Lefevre, S., D.~J. McKenzie, and G.~E. Nilsson. 2018.
\newblock In modelling effects of global warming, invalid assumptions lead to
  unrealistic projections.
\newblock Global change biology 24:553--556.

\bibitem[{Lefrancois and Claireaux(2003)}]{lefrancois_influence_2003}
Lefrancois, C., and G.~Claireaux. 2003.
\newblock Influence of ambient oxygenation and temperature on metabolic scope
  and scope for heart rate in the common sole solea solea.
\newblock Marine Ecology Progress Series 259:273--284.

\bibitem[{Lucas and Priede(1992)}]{lucas_utilization_1992}
Lucas, M.~C., and I.~G. Priede. 1992.
\newblock Utilization of metabolic scope in relation to feeding and activity by
  individual and grouped zebrafish, brachydanio rerio (hamilton-buchanan).
\newblock Journal of Fish Biology 41:175--190.

\bibitem[{Magnuson and {DeStasio}(1997)}]{magnuson_thermal_1997}
Magnuson, J.~J., and B.~T. {DeStasio}. 1997.
\newblock Thermal niche of fishes and global warming.
\newblock Pages 377--408 \emph{in} C.~Wood and D.~{McDonald}, eds. Global
  warming: implications for freshwater and marine fish, vol.~61. Cambridge
  University Press.

\bibitem[{Marty et~al.(2011)Marty, Dieckmann, Rochet, and
  Ernande}]{marty_impact_2011}
Marty, L., U.~Dieckmann, M.-J. Rochet, and B.~Ernande. 2011.
\newblock Impact of environmental covariation in growth and mortality on
  evolving maturation reaction norms.
\newblock The American Naturalist 177:E98--E118.

\bibitem[{Myrick and Cech(2000)}]{myrick_temperature_2000}
Myrick, C.~A., and J.~J. Cech. 2000.
\newblock Temperature influences on california rainbow trout physiological
  performance.
\newblock Fish Physiology and Biochemistry 22:245--254.

\bibitem[{Parmesan and Yohe(2003)}]{parmesan_globally_2003}
Parmesan, C., and G.~Yohe. 2003.
\newblock A globally coherent fingerprint of climate change impacts across
  natural systems.
\newblock Nature 421:37--42.

\bibitem[{Pauly and Cheung(2017)}]{pauly_sound_2017}
Pauly, D., and W.~W.~L. Cheung. 2017.
\newblock Sound physiological knowledge and principles in modeling shrinking of
  fishes under climate change.
\newblock Global Change Biology 24:e15--e26.

\bibitem[{Pinsky et~al.(2013)Pinsky, Worm, Fogarty, Sarmiento, and
  Levin}]{pinsky_marine_2013}
Pinsky, M.~L., B.~Worm, M.~J. Fogarty, J.~L. Sarmiento, and S.~A. Levin. 2013.
\newblock Marine taxa track local climate velocities.
\newblock Science 341:1239--1242.

\bibitem[{Priede(1977)}]{priede_natural_1977}
Priede, I.~G. 1977.
\newblock Natural selection for energetic efficiency and the relationship
  between activity level and mortality.
\newblock Nature 267:610--611.

\bibitem[{Priede(1985)}]{priede_metabolic_1985}
---{}---{}---. 1985.
\newblock Metabolic scope in fishes.
\newblock Pages 33--64 \emph{in} Fish Energetics. Springer.

\bibitem[{Pörtner(2010)}]{portner_oxygen-and_2010}
Pörtner, H.-O. 2010.
\newblock Oxygen-and capacity-limitation of thermal tolerance: a matrix for
  integrating climate-related stressor effects in marine ecosystems.
\newblock The Journal of experimental biology 213:881--893.

\bibitem[{Rall et~al.(2012)Rall, Brose, Hartvig, Kalinkat, Schwarzmüller,
  Vucic-Pestic, and Petchey}]{rall_universal_2012}
Rall, B.~C., U.~Brose, M.~Hartvig, G.~Kalinkat, F.~Schwarzmüller,
  O.~Vucic-Pestic, and O.~L. Petchey. 2012.
\newblock Universal temperature and body-mass scaling of feeding rates.
\newblock Phil. Trans. R. Soc. B 367:2923--2934.

\bibitem[{Sainmont et~al.(2015-09)Sainmont, Andersen, Thygesen, Fiksen, and
  Visser}]{sainmont_effective_2015}
Sainmont, J., K.~H. Andersen, U.~H. Thygesen, ?.~Fiksen, and A.~W. Visser.
  2015-09.
\newblock An effective algorithm for approximating adaptive behavior in
  seasonal environments.
\newblock Ecological Modelling 311:20--30.

\bibitem[{Salin et~al.(2016)Salin, Auer, Anderson, Selman, and
  Metcalfe}]{salin_inadequate_2016}
Salin, K., S.~K. Auer, G.~J. Anderson, C.~Selman, and N.~B. Metcalfe. 2016.
\newblock Inadequate food intake at high temperatures is related to depressed
  mitochondrial respiratory capacity.
\newblock Journal of Experimental Biology 219:1356--1362.

\bibitem[{Sentis et~al.(2013)Sentis, Hemptinne, and
  Brodeur}]{sentis_parsing_2013}
Sentis, A., J.-L. Hemptinne, and J.~Brodeur. 2013.
\newblock Parsing handling time into its components: implications for responses
  to a temperature gradient.
\newblock Ecology 94:1675--1680.

\bibitem[{Sswat et~al.(2018)Sswat, Stiasny, Jutfelt, Riebesell, and
  Clemmesen}]{sswat_growth_2018}
Sswat, M., M.~H. Stiasny, F.~Jutfelt, U.~Riebesell, and C.~Clemmesen.
  2018.
\newblock Growth performance and survival of larval atlantic herring, under the
  combined effects of elevated temperatures and {CO}2.
\newblock {PLOS} {ONE} 13:e0191947.

\bibitem[{Stuart-Smith et~al.(2015)Stuart-Smith, Edgar, Barrett, Kininmonth,
  and Bates}]{stuart-smith_thermal_2015}
Stuart-Smith, R.~D., G.~J. Edgar, N.~S. Barrett, S.~J. Kininmonth, and A.~E.
  Bates. 2015.
\newblock Thermal biases and vulnerability to warming in the world’s marine
  fauna.
\newblock Nature 528:88--92.

\bibitem[{Verberk et~al.(2016)Verberk, Overgaard, Ern, Bayley, Wang, Boardman,
  and Terblanche}]{verberk_does_2016}
Verberk, W. C. E.~P., J.~Overgaard, R.~Ern, M.~Bayley, T.~Wang, L.~Boardman,
  and J.~S. Terblanche. 2016.
\newblock Does oxygen limit thermal tolerance in arthropods? a critical review
  of current evidence.
\newblock Comparative Biochemistry and Physiology Part A: Molecular \&
  Integrative Physiology 192:64--78.

\bibitem[{Vucic-Pestic et~al.(2011-03-01)Vucic-Pestic, Ehnes, Rall, and
  Brose}]{vucic-pestic_warming_2011}
Vucic-Pestic, O., R.~B. Ehnes, B.~C. Rall, and U.~Brose. 2011-03-01.
\newblock Warming up the system: higher predator feeding rates but lower
  energetic efficiencies.
\newblock Global Change Biology 17:1301--1310.

\bibitem[{Walther et~al.(2002)Walther, Post, Convey, Menzel, Parmesan, Beebee,
  Fromentin, Hoegh-Guldberg, and Bairlein}]{walther_ecological_2002}
Walther, G.-R., E.~Post, P.~Convey, A.~Menzel, C.~Parmesan, T.~J.~C. Beebee,
  J.-M. Fromentin, O.~Hoegh-Guldberg, and F.~Bairlein. 2002.
\newblock Ecological responses to recent climate change.
\newblock Nature 416:389--395.

\section*{Tables}
\renewcommand{\thetable}{\arabic{table}}
\setcounter{table}{0}

\begin{table}
{\small
\caption{Parameters of the constrained activity model for two scenarios: slow strategy and fast strategy species. For parameters with dual values (i.e., x/y), the former reflects species with a domed maximum oxygen supply (MOS) with respect to temperature, whereas the latter corresponds to species with a continuously increasing MOS. Values in brackets for $\gamma$ are alternative resource availability scenarios.}
\label{tab:parameters}
\begin{tabular}{llcc}
 \multirow{2}{*}{Description} &  \multirow{2}{*}{Symbol (unit)} & \multicolumn{2}{r}{Value} \\
 \cmidrule{3-4}
 & & slow strategy  & fast strategy \\

\hline
\end{tabular}
}
\end{table}

\section*{Figure legends}

\begin{figure}
%\includegraphics[width=\linewidth]{images/O2_fig_used-1} 
\caption[Maximum oxygen supply relative to the maximum supply for species with a dome-shaped MOS (here $\eta$ = 3) and a continually increasing MOS ($\eta$ = 0.1) used in model scenarios discussed below]{Maximum oxygen supply relative to the maximum supply for species with a dome-shaped MOS (here $\eta$ = 3) and a continually increasing MOS ($\eta$ = 0.1) used in model scenarios discussed below.}\label{fig:O2_fig_used}
\end{figure}

<< setup,echo=FALSE,include=F>>=
knitr::opts_chunk$set(echo = F,warning = F, message = F,cache=T,autodep = T)

@@

lab1 <- as_labeller(c(`M strategy Domed`  ='a',
                   `P strategy Domed` = 'b'))

lab2 <- as_labeller(c(`M strategy Domed`  ='c',
                   `P strategy Domed` = 'd'))
